A Type System for String Sanitation\\Implemented Inside a Python

### 2014-08-05 sanitation-psp14/sanitation-psp14-short.tex
Fixed typos.

%\category{D.3.2}{Programming Languages}{Language Classifications}[Extensible Languages]
%\category{D.3.4}{Programming Languages}{Processors}[Compilers]
%\category{F.3.1}{Logics \& Meanings of Programs}{Specifying and Verifying and Reasoning about Programs}[Specification Techniques]
%\keywords
%type-level computation, typed compilation
\section{Introduction}\label{intro}
Improper input sanitation is a leading cause of security vulnerabilities in web applications [OWASP]. Command injection attacks exploit improper input sanitation by inserting malicious code into an otherwise benign command. Modern web frameworks, libraries, and database abstraction layers attempt to ensure proper sanitation of user input. When these methods are unavailable or insufficient, developers implement custom sanitation techniques. Im both cases, sanitation algorithms are implemented using the language's regular expression capabilities and usually \emph{replace} potentially unsafe strings with equivalent escaped strings.

@@

To demonstrate this approach, we present a simply typed lambda calculus with \emph{constrained strings}; that is, a set of string types parameterized by regular expressions. If $s : \stringin{r}$, then $s$ is a string matching the language $r$. Additionally, we include an operation $\rreplace{r}{s_1}{s_2}$ which corresponds to the replace mechanism available in most regular expression libraries; that is, any substring of $s_1$ matching $r$ is replaced with $s_2$. The type of this expression is the computed, and is likely ``smaller" or more constrained than the type of $s_1$. Libraries, frameworks or functions which construct and execute commands containing input can specify a safe subset $\stringin{r_{\sf spec}}$ of strings, and input sanitation algorithms can construct such a string using ${\sf rreplace}$ or, optionally, by coercion (in which case a runtime check is inserted). We also show how this simple can be translated into a host language containing a regular expression library such that the safety guarantee of the extended language is preserved.

@@

Unfortunately, we are unable to present full proofs here due to space constraints.
%Note that we never insert an internal check where the type of the string implies
%that a check must succeed. Furthermore, expensive calculations (such as language
%inclusion or ${\sf rreplace}$) all occur at compile time.
%Since sanitation problems are generally expressible with smalle, simple expressions,
%we believe that the compile-time overhead is not significant enough to prohibiour target usecase. Informal experimentation with our implementation seems to
%support this assertion.

@@

\begin{proof}
The proof proceeds by induction on the typing relation for $S$ and an appropriate
choice of $P$; in each case, the choice is obvious. The subcases (a) proceed
by inversion and appeals to the induction hypothesis. The subcases (b) proceed
by the standard error propagation rules omitted for space.
Throughout the proof, properties from the closure lemma for regular languages
are necessary.

@@

Composable analyses which complement existing approaches constitute a promising approach toward the integration of security concerns into programming languages.
In this paper, we presented a system with both of these properties and defined a security-prserving transformation.
Unlike other approaches, our solution complements existing, familiar solutions while providing a strong guarantee that traditiona library and framework-based approaches are implemented and utilized correctly.

>>>>

%\category{D.3.2}{Programming Languages}{Language Classifications}[Extensible Languages]
%\category{D.3.4}{Programming Languages}{Processors}[Compilers]
%\category{F.3.1}{Logics \& Meanings of Programs}{Specifying and Verifying and Reasoning about Programs}[Specification Techniques]
%\keywords
%type-level computation, typed compilation
\section{Introduction}\label{intro}
Improper input sanitation is a leading cause of security vulnerabilities in web applications [OWASP]. Command injection attacks exploit improper input sanitation by inserting malicious code into an otherwise benign command. Modern web frameworks, libraries, and database abstraction layers attempt to ensure proper sanitation of user input. When these methods are unavailable or insufficient, developers implement custom sanitation techniques. In both cases, sanitation algorithms are implemented using the language's regular expression capabilities and usually \emph{replace} potentially unsafe strings with equivalent escaped strings.

@@

To demonstrate this approach, we present a simply typed lambda calculus with \emph{constrained strings}; that is, a set of string types parameterized by regular expressions. If $s : \stringin{r}$, then $s$ is a string matching the language $r$. Additionally, we include an operation $\rreplace{r}{s_1}{s_2}$ which corresponds to the replace mechanism available in most regular expression libraries; that is, any substring of $s_1$ matching $r$ is replaced with $s_2$. The type of this expression is the computed, and is likely ``smaller" or more constrained than the type of $s_1$. Libraries, frameworks or functions which construct and execute commands containing input can specify a safe subset $\stringin{r_{\sf spec}}$ of strings, and input sanitation algorithms can construct such a string using ${\sf rreplace}$ or, optionally, by coercion (in which case a runtime check is inserted). We also show how this system is translated into a host language containing a regular expression library such that the safety guarantee of the extended language is preserved.

@@

Unfortunately, we are unable to present full proofs in this paper due to space constraints; however, proofs are given in full in the accompanying technical report?.
%Note that we never insert an internal check where the type of the string implies
%that a check must succeed. Furthermore, expensive calculations (such as language
%inclusion or ${\sf rreplace}$) all occur at compile time.
%Since sanitation problems are generally expressible with smalle, simple expressions,
%we believe that the compile-time overhead is not significant enough to prohibiour target usecase. Informal experimentation with our implementation seems to
%support this assertion.

@@

\begin{proof}
The proof proceeds by induction on the typing relation for $S$ and an appropriate
choice of $P$; in each case, the choice is obvious. The subcases (a) proceed
by inversion and appeals to our type safety theorems as well as the induction hypothesis. The subcases (b) proceed
by the standard error propagation rules omitted for space.
Throughout the proof, properties from the closure lemma for regular languages
are necessary.

@@

Composable analyses which complement existing approaches constitute a promising approach toward the integration of security concerns into programming languages.
In this paper, we presented a system with both of these properties and defined a security-preserving transformation.
Unlike other approaches, our solution complements existing, familiar solutions while providing a strong guarantee that traditional library and framework-based approaches are implemented and utilized correctly.

### 2014-08-05 sanitation-psp14/sanitation-psp14.tex
Fixed typos.

To demonstrate this approach, we present a simply typed lambda calculus with \emph{constrained strings}; that is, a set of string types parameterized by regular expressions. If $s : \stringin{r}$, then $s$ is a string matching the language $r$. Additionally, we include an operation $\rreplace{r}{s_1}{s_2}$ which corresponds to the replace mechanism available in most regular expression libraries; that is, any substring of $s_1$ matching $r$ is replaced with $s_2$. The type of this expression is the computed, and is likely ``smaller" or more constrained than the type of $s_1$. Libraries, frameworks or functions which construct and execute commands containing input can specify a safe subset $\stringin{r_{\sf spec}}$ of strings, and input sanitation algorithms can construct such a string using ${\sf rreplace}$ or, optionally, by coercion (in which case a runtime check is inserted). We also show how this simple can be translated into a host language containing a regular expression library such that the safety guarantee of the extended language is preserved.

>>>>

To demonstrate this approach, we present a simply typed lambda calculus with \emph{constrained strings}; that is, a set of string types parameterized by regular expressions. If $s : \stringin{r}$, then $s$ is a string matching the language $r$. Additionally, we include an operation $\rreplace{r}{s_1}{s_2}$ which corresponds to the replace mechanism available in most regular expression libraries; that is, any substring of $s_1$ matching $r$ is replaced with $s_2$. The type of this expression is computed, and is likely ``smaller" or more constrained than the type of $s_1$. Libraries, frameworks or functions which construct and execute commands containing input can specify a safe subset $\stringin{r_{\sf spec}}$ of strings, and input sanitation algorithms can construct a string with this type using ${\sf rreplace}$ or, optionally, by coercion (in which case a runtime check is inserted). We also show how this simple extension can be translated into a host language containing a regular expression library such that the safety guarantee of the extended language is preserved.
